% RQuantLib examples
% http://dirk.eddelbuettel.com/code/rquantlib.html

European Option Example
=======================

## A simple example for EuropeanOption
Compute price using Black-Scholes formula

* type = ``call` - put or call
* S0 (underlying) = 100
* K (strike) = 100
* c (dividendYield) = 0.01
* r (riskFreeRate) = 0.03
* T (maturity) = 0.5
* sigma (volatility) = 0.5

```{r}
library(RQuantLib)

eo <- EuropeanOption("call", 100, 100, 0.01, 0.03, 0.5, 0.4)
print(eo)
summary(eo)
```

## Computing implied volatility

```{r}
eo.vol <- EuropeanOptionImpliedVolatility("call", value=eo$value+0.5, 100, 100, 0.01, 0.03, 0.5, 0.4)
print(eo.vol)
print(eo.vol$impliedVol)
```

## Vectorized computation
The above function only deals with scalars. Below we can (exactly two) input vectors and get vectors as output, e.g. for a range of S0 and sigma.

```{r}
S0 <- seq(10, 180, by=5)
sigma <- seq(0.2, 0.8, by=0.1)
#T0 <- seq(0.8, 1.2, by=0.1)

eo.price <- EuropeanOptionArrays("call", underlying=S0,
                                 strike=100,
                                 dividendYield=0.01,
                                 riskFreeRate=0.03,
                                 maturity=1,
                                 volatility=sigma)                               


dim(eo.price$value)   # == c(length(S0), length(sigma))
```


Plot using ggplot2

```{r}
library(reshape2)
library(ggplot2)

# plot S0 vs (value, greeks). different lines for different sigma
eo.df <- data.frame(S0=S0,
                    sigma=rep(sigma,each=length(S0)),
                    value=as.numeric(eo.price$value),  # column-major, ok
                    delta=as.numeric(eo.price$delta),
                    gamma=as.numeric(eo.price$gamma),
                    vega=as.numeric(eo.price$vega))

eo.df$sigma <- as.factor(eo.df$sigma)


eo.df2 <- melt(eo.df, id=c("S0", "sigma"))
#qplot(S0, value, data=eo.df2, colour=sigma, geom="line") + facet_wrap(~variable, ncol=2)

#old.par <- par()
#par(mfrow=c(2,2), oma=c(5,0,0,0), mar=c(2,2,2,1))
par(old.par)

qplot(S0, value, data=subset(eo.df2,variable=="value"), colour=sigma, geom="line") +
  labs(title="Price", y="P")

qplot(S0, value, data=subset(eo.df2,variable=="delta"), colour=sigma, geom="line") +
  labs(title="Delta", y=expression(delta))

qplot(S0, value, data=subset(eo.df2,variable=="gamma"), colour=sigma, geom="line") +
  labs(title="Gamma", y=expression(gamma))

qplot(S0, value, data=subset(eo.df2,variable=="vega"), colour=sigma, geom="line") +
  labs(title="Vega", y=expression(vega))
```


Discount Curve Example
======================

See ``example(DiscountCurve)``

```{r}
params <- list(tradeDate=as.Date('2004-09-20'),
               settleDate=as.Date('2004-09-22'),
               dt=.25,
               interpWhat="discount",
               interpHow="loglinear")

# Settings::instance().evaluationDate() = Date(22, Nov, 2004);
setEvaluationDate(as.Date("2004-11-22"))

times <- seq(0, 10, .1)

tsQuote1 <- list(d1w = 0.0382,          # 1-week deposit rate
                 d1m = 0.0372,          # 1-month deposit rate
                 d3m = 0.0363,          # ...
                 d6m = 0.0358,
                 d9m = 0.0348,
                 d1y = 0.0345,          # 1-yeardeposit rate
                 #fut1 = 96.2875,
                 #fut2 = 96.7875,
                 #fut3 = 96.9875,
                 #fut4 = 96.6875,
                 #fut5 = 96.4875,
                 #fut6 = 96.3875,
                 #fut7 = 96.2875,
                 #fut8 = 96.0875,
                 s2y = 0.037125,        # 2-year swap rate
                 s3y = 0.0398,
                 s5y = 0.0443,
                 s10y = 0.05165,
                 s15y = 0.055175)

# error here
#curve1 <- DiscountCurve(params, tsQuote1, times)
#plot(curve1, setpar=FALSE)

tsQuote2 <- list("flat" = 0.02)
curve2 <- DiscountCurve(params, tsQuote2, times)
plot(curve2, setpar=FALSE)
```

Example from site

```{r}
params <- list(tradeDate = c(2, 15, 2002),
               settleDate = c(2, 19, 2002),
               dt = 0.25,
               interpWhat = "discount",
               interpHow = "loglinear")

setEvaluationDate(as.Date("2002-02-19"))
tsQuote3 <- list(d1w = 0.0382, d1m = 0.0372, fut1 = 96.2875,
    fut2 = 96.7875, fut3 = 96.9875, fut4 = 96.6875, fut5 = 96.4875,
    fut6 = 96.3875, fut7 = 96.2875, fut8 = 96.0875, s3y = 0.0398,
    s5y = 0.0443, s10y = 0.05165, s15y = 0.055175)

times <- seq(0, 10, 0.1)

savepar <- par(mfrow=c(3,3))

curves <- DiscountCurve(params, tsQuote3, times)
plot(curves, setpar=FALSE)

params$interpHow <- "linear"
curves <- DiscountCurve(params, tsQuote3, times)
plot(curves, setpar=FALSE)

params$interpHow <- "spline"
curves <- DiscountCurve(params, tsQuote3, times)
plot(curves, setpar=FALSE)

```


